A weakly structured stem for human origins in Africa

### 2022-01-07 paper.tex
suggestion

Regardless of the model choice for early epochs, inference of recent human
demographic history from two thousand \aprcomment{why two thousand, instead of
just ``up to 150kya''?} to 150kya was remarkably robust. The earliest divergence
among contemporary sampled human populations differentiates the southern
African Nama from other African groups between 110--135kya, with low to moderate
levels of subsequent gene flow \aprcomment{(Table 1)}. In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim$150kya. We conclude that, regardless of model
specification, geographic patterns of \emph{Homo sapiens} population structure
date back to the Late Middle Stone Age in Africa, likely arising during MIS 5.
Although we find evidence for earlier population structure in Africa (see
below), present-day populations cannot be easily mapped onto these more ancient
stem groups as only a small proportion of drift between modern populations can
be attributed to drift between stems \aprcomment{(Figure 3)}.

@@

Likelihood-based parameter confidence intervals are relatively narrow within
each specified model
(Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}),
reflecting an informative statistical approach. As expected, model assumptions
have a larger impact on parameter estimates. To convey model uncertainty, we
highlight features of two models with the highest \aprcomment{not the
highest\ldots merger-without-stem-migration at -107,650}log likelihoods. These
differ primarily in whether early stems are assumed to give rise to modern
populations in multiple discrete merger events or in continuous migrations
\aprcomment{(Figure 2)}. Additional models are discussed in
\aprcomment{Supplement XXX}.

>>>>

Regardless of the model choice for early epochs, inference of human
demographic history in the last 150kya was remarkably robust. The earliest divergence
among contemporary sampled human populations differentiates the southern
African Nama from other African groups between 110--135kya, with low to moderate
levels of subsequent gene flow \aprcomment{(Table 1)}. In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim$150kya. We conclude that, regardless of model
specification, geographic patterns of \emph{Homo sapiens} population structure
date back to the Late Middle Stone Age in Africa, likely arising during MIS 5.
Although we find evidence for earlier population structure in Africa (see
below), present-day populations cannot be easily mapped onto these more ancient
stem groups as only a small proportion of drift between modern populations can
be attributed to drift between stems \aprcomment{(Figure 3)}.

@@

Likelihood-based parameter confidence intervals  \sgcomment{or: Bootstrap-based confidence interval reflecting the finite length of the genome?}   are relatively narrow within
each specified model
(Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}),
reflecting an informative statistical approach. As expected, model assumptions
have a larger impact on parameter estimates. To convey model uncertainty, we
highlight features of two models with the highest \aprcomment{not the
highest\ldots merger-without-stem-migration at -107,650}log likelihoods. These
differ primarily in whether early stems are assumed to give rise to modern
populations in multiple discrete merger events or in continuous migrations
\aprcomment{(Figure 2)}. Additional models are discussed in
\aprcomment{Supplement XXX}.
